---
title: "MapOSR"
subtitle: "Inter-Rater-Agreement"
author: "Jürgen Schneider"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

# Setup

Libraries and R version.

```{r setup, warning=F, message=F, comment=F}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE, 
                      comment=FALSE)
library(tidyverse)
library(irr)
library(kableExtra)
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
R.Version()
```

# Import & Wrangling

Data import. You can get the underlying data for this document on github ([github.com/JanaLasser/mapping-open-science-research](https://github.com/JanaLasser/mapping-open-science-research)) or `r xfun::embed_file(here::here("02_data_tidy/reliability_clean.csv"), text="download it directly from this document")`.

```{r import}
# IMPORT ########################################################################
rel <- rio::import(here::here("02_data_tidy/reliability_clean.csv"))
```

There were 5 categories to be coded:

* action
* method
* discipline
* group
* geo

Since the categories had multiple-choice answer labels, these labels were dichotomized (dummy coded). For each __label__ on these categories, there is now a variable (e.g. group_librarian, group_policy, ...) on which "1" (applies to the publication) or "0" (does _not_ apply to the publication) was coded.

Variables beginning with `c1_` refer to the first coder, variables beginning with `c2_` refer to the second coder.  

Accordingly, this type of table:

```{r import_out1}
rel[1:5,] %>%
  kbl() %>%
  kable_styling()
```

\
\
\

will be transformed through this code (click on 'code' button to see):

```{r wrangling}
# WRANGLING #####################################################################
rel <- rel %>%
  mutate(across(2:11, ~ str_remove_all(., "(a=|m=|d=|t=|geo=)"))) %>% # delete all prefix
  pivot_longer(2:11, names_to = "variable", values_to = "values") %>%
  separate_rows(values) %>%                  # multiple entries in cells as multiple rows
  mutate(values = trimws(tolower(values)),   # delete whitespace, lowercase
         values = case_when(
                      values == "librarians" ~ "librarian",   # Typo recoding
                      values == "bibliometric" ~ "biblio",    # there is no geo code zar -> zaf?
                      values == "openpolicy" ~ "openpolicies",
                      values == "opentool" ~ "opentools",
                      values == "publishers" ~ "publisher",
                      values == "researchers" ~ "researcher",
                      values == "soscie" ~ "socscie",
                      str_detect(variable, "geo") & (values == "none" | values == "all") ~ "unspecific",
                      str_detect(variable, "discipline") & (values == "none" | values == "all") ~ "unspecific",
                      TRUE ~ as.character(values)),
         cells = 1) %>%                             # add new variable for coded values
  dplyr::filter(values != "" & !is.na(values)) %>%  # delete empty cells
  pivot_wider(id_cols = ID, 
              names_from = c(variable, values),
              values_from = cells, 
              values_fill = 0)                      # add 0 indicating "not coded"
```

\
\
\


into this kind of table:

```{r import_out2}
rel[1:5,] %>%
  kbl() %>%
  kable_styling()
```

# Agreement

## No Agreement

Are there any answer categories that __weren't__ given by one person (but were by the other)?

```{r noAgreement}
# AGREEMENT ###################################################################
# checking which variables exist in c1 and c2 (and are thus coded by both coders)
kappa_variablen <- data.frame(variablen = str_remove(names(rel[-1]), "(c1_|c2_)")) 

kappa_variablen <- kappa_variablen %>%
                      group_by(variablen) %>%
                      dplyr::summarise(count = n())

# Variables WITHOUR agreement
kappa_variablen %>%
  dplyr::filter(count != 2) %>%
  print(n=30)
```

## Computing Agreement

Loop over all response categories coded by both coders.

```{r compAgreement}
# Variables for which you can calculate match
kappa_variablen <- kappa_variablen %>%
  dplyr::filter(count == 2)

# generate empty data frame for results
kappa_results <- data.frame(variable = as.character(),
                            kappa    = as.numeric(),
                            percent  = as.numeric())

#loop over all dichotomous variables
for(i in kappa_variablen$variablen) {
  
  tmp_kappa <- kappa2(data.frame(rel[paste0("c1_", i)],
                                 rel[paste0("c2_", i)]))
  
  tmp_perc <- agree(data.frame(rel[paste0("c1_", i)],
                               rel[paste0("c2_", i)]))
  
  kappa_results <- kappa_results %>%
    add_row(variable = i,
            kappa    = tmp_kappa$value,
            percent  = tmp_perc$value)
}
```

## Agrement Results: Kappa

Visualization and table of Cohen's kappa for each dichotomous variable.

```{r kappa}
# plot results kappa
ggplot(kappa_results, aes(x=kappa)) +
  geom_density(size=2) +
  geom_dotplot() +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1.1)) +
  geom_vline(xintercept = .7, color="red", alpha=.5, linetype="dashed", size=1) +
  theme_light()

# results table
# ordered by kappa
kappa_results[order(kappa_results$kappa),] %>%
  kbl() %>%
  kable_styling()
```

## Agrement Results: Percentage

Since the dichotomous response categories are sometimes distributed very skewed, even a few deviations lead to very small kappa values. Percentage matches are calculated accordingly.

```{r perc}
# plot results percent
ggplot(kappa_results, aes(x=percent)) +
  geom_density(size=1) +
  geom_dotplot(binwidth = 1, size=2) +
  scale_x_continuous(expand=c(0,0), limits = c(0,100)) +
  scale_y_continuous(expand=c(0,0), limits = c(0,.075)) +
  geom_vline(xintercept = 80, color="red", alpha=.5, linetype="dashed", size=1) +
  theme_light()

# results table
# ordered by percent
kappa_results[order(kappa_results$percent),] %>%
  kbl() %>%
  kable_styling()
```

## Agrement Results: Summary


```{r summary1}
# summary of results
summary(kappa_results[2:3])
```

## Agrement Results: Summary grouped

\

__Kappa:__
```{r summary2}
# summary of results
kappa_results %>%
  group_by(str_sub(variable, 1, 3)) %>%
  summarise(min = min(kappa),
            max = max(kappa),
            median = median(kappa)) %>%
  kbl() %>%
  kable_styling()
```
\
\

__Percent:__
```{r summary3}
# summary of results
kappa_results %>%
  group_by(str_sub(variable, 1, 3)) %>%
  summarise(min = min(percent),
            max = max(percent),
            median = median(percent)) %>%
  kbl() %>%
  kable_styling()
```